---
title: "lclapp_ME4320_Player3"
author: "Lewis Clapp"
date: "2025-12-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
library(agricolae)
```

Read csv
```{r}
Player3<-read.csv("Player3_Data.csv",header=TRUE)
```

Check normallity of Speed with shapiro
```{r}
with(Player3,shapiro.test(Speed[Treatment=="A"]))

with(Player3,shapiro.test(Speed[Treatment=="B"]))

with(Player3,shapiro.test(Speed[Treatment=="C"]))

```

Check  Equal variance of speed with Lavenes because normallity could not be assumed
```{r}
with(Player3,leveneTest(Speed~Treatment))
```
Kruskal Wallis with speed because assumptions could not be met
```{r}
kruskal.test(Speed~Treatment,data=Player3)
```
LSD to see which one is different for speed
```{r}
Players3Speed_aov<-aov(Speed~Treatment,data=Player3)
 print(LSD.test(Players3Speed_aov,trt="Treatment",p.adj="bonferroni"))
```



Check normallity of rotation with shapiro
```{r}
with(Player3,shapiro.test(Rotation[Treatment=="A"]))

with(Player3,shapiro.test(Rotation[Treatment=="B"]))

with(Player3,shapiro.test(Rotation[Treatment=="C"]))

```

Check  Equal variance of Rotation with Bartlett
```{r}
with(Player3,bartlett.test(Rotation~Treatment))
```


ANOVA with speed because assumptions were verified
```{r}
Player3Rotation_aov<-aov(Rotation~Treatment,data=Player3)
 summary(Player3Rotation_aov)
```

Tukeys to see which Rotation one is different
```{r}
TukeyHSD(Player3Rotation_aov,conf.level=0.95)
```







Check Normality with Energy with Shapiro
```{r}
with(Player3,shapiro.test(Energy[Treatment=="A"]))

with(Player3,shapiro.test(Energy[Treatment=="B"]))

with(Player3,shapiro.test(Energy[Treatment=="C"]))

```
Check  Equal variance of Energy with Bartlett
```{r}
with(Player3,bartlett.test(Energy~Treatment))
```
ANOVA with Energy because assumptions were verified
```{r}
Player3Energy_aov<-aov(Energy~Treatment,data=Player3)
 summary(Player3Energy_aov)
```
Tukeys to see which Energy one is different
```{r}
TukeyHSD(Player3Energy_aov,conf.level=0.95)
```